Exploring probability distributions for bivariate temporal granularities

### 2019-07-28 main.Rmd
adding brief on visualisation and harmony/clash

Consider a hierarchy table laying out the structure of all temporal granularities we are interested to look at. The hierarchy table consists of three columns: 
 - The first column represents the temporal granularities in ascending order. 
 - The second column represents the constant which relates subsequent granularities.
 - The third column provides the functional form for single-order up-granularities

@@

\begin{equation} \label{eq1}
\begin{split}
f(x, y) & = f(x, x+1) + c(x, x+1)*[f(x+1, y)-1] \\
  & =  f(x, x+1) + c(x, x+1)[f(x+1, x+2) + c(x+1, x+2)[f(x+2, y) - 1] - 1]\\
  & =  f(x, x+1) + c(x, x+1)[f(x+1, x+2) - 1] + c(x, x+1)c(x+1, x+2)[f(x+2, y) - 1]
\end{split}
\end{equation}

@@

@@

    We could also derive f(w, y) given f(x, y) and the hierarchy table, for $order(y)>=order(w) > order(x)$.

@@

@@

>>>>

Consider a hierarchy table laying out the structure of all temporal granularities we are interested to look at, consisting of the following  three columns: 
 - The first column represents the temporal granularities in ascending order.   
 - The second column represents the constant which relates subsequent granularities.  
 - The third column provides the functional form for single-order up-granularities.

@@

\begin{equation} \label{eq1}
\begin{split}
f(x, y) & = f(x, x+1) + c(x, x+1)*[f(x+1, y)-1] \\  
  & =  f(x, x+1) + c(x, x+1)[f(x+1, x+2) + c(x+1, x+2)[f(x+2, y) - 1] - 1]\\ (expanding f(x+1, y))  
  & =  f(x, x+1) + c(x, x+1)[f(x+1, x+2) - 1] + c(x, x+1)c(x+1, x+2)[f(x+2, y) - 1]
\end{split}
\end{equation}

@@

Suppose, we are not given the intermediate single-order-up granularities but still want all possible multiple order-up granularties. In that case, it is necessary to atleast have the have the functional form that relates the smallest temporal unit to the highest temporal unit. In that case, we are able to form single-order-up from this functional relationship and still use the recursive relation to get the multiple order-up granularities.

@@

  We could also derive f(w, y) given f(x, y) and the hierarchy table, for $order(y)>=order(w) > order(x).

@@

We investigate some combinations of circular time granularities which facilitate or hinder exploratory analysis. The combinations of circular granularities which promote the exploratory analysis through visualization are referred to as **harmonies** and the ones which impede the analysis are referred to as **clashes**.

Let's take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month. Here $C_1$ can take 31 values while $C_2$ can take 5 values. There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM. Many of these are empty. For example $S_{1,5}$, $S_{21,2}$, etc. In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful. These are structurally empty sets in that it is impossible for them to have any observations.

Another example could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year. Here $C_1$ can take 7 values while $C_2$ can take 12 values. So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY. All of these are non-empty because every DOW can occur in every month. So graphics involving $C_1$ and $C_2$ are potentially useful.

@@

With huge amount of data currently available, it is important to visualize the distribution of the time series variable rather than any particular statistic. But which ditribution plot to draw, is based on what features of the distribution we are interested to look at, or how much display space is available to us.

Offcourse, the time series variable can be plotted against many time granularities to get more understanding of the underlying periodicty, however, we will restrict ourselves to see the distribution of the time series across bivariate temporal granularities. That neccessiates plotting temporal granularity along the x-axis and the other one across facets.

Now, due to the hierarchical arrangement of the granularities, there are certain granularities which when plotted together do not give us the layout to do exploration.
The harmonies when plotted together can help exploration, however, the clashes were plotted against each other is not appropriate for exploration.

We investigate some combinations of circular time granularities which facilitate or hinder exploratory analysis. Since we are restricting ourselves to 2D space, we assume that the circular granularities are plotted across the aesthetics and facets. The combinations of circular granularities which promote the exploratory analysis through visualization are referred to as **harmonies** and the ones which impede the analysis are referred to as **clashes**.
